---
title: "Correlation"
author: "Kelly Chin"
date: "3/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R libraries
```{r library, message=FALSE, warning=FALSE}
library(readxl)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(reader)
library(lubridate)
library(docstring)
library(roxygen2)
library(corrplot)
``` 


# Correlation in farm info, ketosis, reproduction, milk yield
```{r}
df.farm <- Three %>%
  select(FarmNo, CowId, CowNo, Lactation, ad_Lactation)
str(df.farm)
df.farm$FarmNo <- as.integer(df.farm$FarmNo)
cor.farm <- cor(df.farm)
cor.farm
corrplot(cor.farm , tl.col = "brown",
         tl.srt = 45, bg = "White",
         title = "\n\n Correlation plot of farm info",
         addCoef.col = "black",
         type = "lower")

df.repro <- Three %>%
  select(CowId, DFC_CLA, DFCFirstHeat, DaysToFirstIns, Last_AI, ad_Lactation)
str(df.repro)
cor.repro <- cor(na.omit(df.repro))
cor.repro
corrplot(cor.repro,
         tl.col = "brown",
         tl.srt = 45, bg = "white",
         title = "\n\n Correlation plot of reproduction parameters",
         addCoef.col = "black",
         type = "lower")

df.repro$ad_Lactation <- as.factor(df.repro$ad_Lactation)
pairs(df.repro, 
      labels = colnames(df.repro),
      pch = 21, 
      bg = rainbow(4)[df.repro$ad_Lactation],
      col = rainbow(4)[df.repro$ad_Lactation],
      main = "df.repro",
      rowlattop = TRUE,
      gap = 1,
      cex.labels = NULL,
      font.labels = 1)

df.keto20 <- Three %>%
  select(CowId, AvgKeto20_atleast10, KetoDur20, MaxKetoDFC20, nHighKeto20)
cor.keto20 <- cor(na.omit(df.keto20))
cor.keto20
corrplot(cor.keto20,
         tl.col = "brown",
         tl.srt = 45, bg = "white",
         title = "\n\n Correlation plot of ketosis parameters",
         addCoef.col = "black",
         type = "lower")

df.keto60 <- Three %>%
  select(CowId, AvgKeto60_atleast10, KetoDur60, MaxKetoDFC60, nHighKeto60)
cor.keto60 <- cor(na.omit(df.keto60))
cor.keto60
corrplot(cor.keto60,
         tl.col = "brown",
         tl.srt = 45, bg = "white",
         title = "\n\n Correlation plot of ketosis parameters",
         addCoef.col = "black",
         type = "lower")

df.milk <- Three %>%
  select(CowId, PeakYield, PeakDim, AvgMY60, SumMY60, AvgMY150, SumMY150)
cor.milk60 <- cor(na.omit(df.milk))
cor.milk60
corrplot(cor.milk60,
         tl.col = "brown",
         tl.srt = 45, bg = "white",
         title = "\n\n Correlation plot of milk parameters",
         addCoef.col = "black",
         type = "lower")
```


# Reproduction parameters
```{r}
plot(Three$DFC_CLA, Three$DFCFirstHeat, 
     main = "Scatterplot of DFC to first cyclicity and first heat", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first heat", pch = 19)

plot(Three$DFCFirstHeat, Three$DaysToFirstIns, 
     main = "Scatterplot of DFC to first heat and first AI", 
     xlab = "DFC to frist heat", 
     ylab = "DFC to first AI", pch = 19)

plot(Three$DFC_CLA, Three$DaysToFirstIns, 
     main = "Scatterplot of DFC to first cyclicity and first AI", 
     xlab = "DFC to frist cyclicity", 
     ylab = "DFC to first AI", pch = 19)

plot(Three$ad_Lactation, Three$DFC_CLA,
     main = "Scatterplot of Lactation and DFC to first cyclicity",
     xlab = "Adjusted lacation groups",
     ylab = "DFC to first cyclicity", pch = 19)
```
```{r}
# Split data with different lactation groups
df.parity1 <- Three %>%
  filter(ad_Lactation == 1)
df.parity2 <- Three %>%
  filter(ad_Lactation == 2)
df.parity3 <- Three %>%
  filter(ad_Lactation == 3)
df.parity4 <- Three %>%
  filter(ad_Lactation == 4)
```
## Reproduction parameters - DFC_CAL vs DFCFirstHeat
```{r}
par(mfrow = c(1, 4))
plot(df.parity1$DFC_CLA, df.parity1$DFCFirstHeat, 
     #main = "DFC to first cyclicity and first heat in first lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first heat (lactation = 1)", pch = 19)
abline(lm(df.parity1$DFCFirstHeat~df.parity1$DFC_CLA))

plot(df.parity2$DFC_CLA, df.parity2$DFCFirstHeat, 
     #main = "DFC to first cyclicity and first heat in second lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first heat (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DFCFirstHeat~df.parity2$DFC_CLA), col = "red")

plot(df.parity3$DFC_CLA, df.parity3$DFCFirstHeat, 
     #main = "DFC to first cyclicity and first heat in third lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first heat (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DFCFirstHeat~df.parity3$DFC_CLA), col = "green")

plot(df.parity4$DFC_CLA, df.parity4$DFCFirstHeat, 
     #main = "DFC to first cyclicity and first heat in third lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first heat (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DFCFirstHeat~df.parity4$DFC_CLA), col = "blue")

## Pearson correlation
cor(df.parity1$DFCFirstHeat, df.parity1$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity2$DFCFirstHeat, df.parity2$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity3$DFCFirstHeat, df.parity3$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity4$DFCFirstHeat, df.parity4$DFC_CLA, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$DFCFirstHeat, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity2$DFCFirstHeat, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity3$DFCFirstHeat, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity4$DFCFirstHeat, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

cor.test(df.parity1$DFCFirstHeat, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity2$DFCFirstHeat, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity3$DFCFirstHeat, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity4$DFCFirstHeat, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")
```

## Reproduction parameters - DFC_CAL vs DaysToFirstIns
```{r}
par(mfrow = c(1, 4))
plot(df.parity1$DFC_CLA, df.parity1$DaysToFirstIns, 
     #main = "DFC to first cyclicity and first AI in first lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first AI (lactation = 1)", pch = 19)
abline(lm(df.parity1$DaysToFirstIns~df.parity1$DFC_CLA))

plot(df.parity2$DFC_CLA, df.parity2$DaysToFirstIns, 
     #main = "DFC to first cyclicity and first AI in second lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first AI (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DaysToFirstIns~df.parity2$DFC_CLA), col = "red")

plot(df.parity3$DFC_CLA, df.parity3$DaysToFirstIns, 
     #main = "DFC to first cyclicity and first AI in third lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first AI (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DaysToFirstIns~df.parity3$DFC_CLA), col = "green")

plot(df.parity4$DFC_CLA, df.parity4$DaysToFirstIns, 
     #main = "DFC to first cyclicity and first AI in third lactation", 
     xlab = "DFC to first cyclicity",
     ylab = "DFC to first AI (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DaysToFirstIns~df.parity4$DFC_CLA), col = "blue")

## Pearson correlation
cor(df.parity1$DaysToFirstIns, df.parity1$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity2$DaysToFirstIns, df.parity2$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity3$DaysToFirstIns, df.parity3$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity4$DaysToFirstIns, df.parity4$DFC_CLA, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$DaysToFirstIns, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity2$DaysToFirstIns, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity3$DaysToFirstIns, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity4$DaysToFirstIns, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

cor.test(df.parity1$DaysToFirstIns, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity2$DaysToFirstIns, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity3$DaysToFirstIns, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity4$DaysToFirstIns, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")
```

## Reproduction parameters - DFCFirstHeat vs DaysToFirstIns
```{r}
plot(Three$DFCFirstHeat, Three$DaysToFirstIns, 
     main = "Scatterplot of DFC to first heat and first AI", 
     xlab = "DFC to frist heat", 
     ylab = "DFC to first AI", pch = 19)

par(mfrow = c(1, 4))
plot(df.parity1$DFCFirstHeat, df.parity1$DaysToFirstIns, 
     #main = "DFC to first heat and first AI", 
     xlab = "DFC to frist heat",
     ylab = "DFC to first AI (lactation = 1)", pch = 19)
abline(lm(df.parity1$DaysToFirstIns~df.parity1$DFCFirstHeat))

plot(df.parity2$DFCFirstHeat, df.parity2$DaysToFirstIns, 
     #main = "DFC to first heat and first AI", 
     xlab = "DFC to frist heat",
     ylab = "DFC to first AI (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DaysToFirstIns~df.parity2$DFCFirstHeat), col = "red")

plot(df.parity3$DFCFirstHeat, df.parity3$DaysToFirstIns, 
     #main = "DFC to first heat and first AI", 
     xlab = "DFC to frist heat",
     ylab = "DFC to first AI (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DaysToFirstIns~df.parity3$DFCFirstHeat), col = "green")

plot(df.parity4$DFCFirstHeat, df.parity4$DaysToFirstIns, 
     #main = "DFC to first heat and first AI", 
     xlab = "DFC to frist heat",
     ylab = "DFC to first AI (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DaysToFirstIns~df.parity4$DFCFirstHeat), col = "blue")

## Pearson correlation
cor(df.parity1$DaysToFirstIns, df.parity1$DFCFirstHeat, use = "complete.obs", method = "pearson")
cor(df.parity2$DaysToFirstIns, df.parity2$DFCFirstHeat, use = "complete.obs", method = "pearson")
cor(df.parity3$DaysToFirstIns, df.parity3$DFCFirstHeat, use = "complete.obs", method = "pearson")
cor(df.parity4$DaysToFirstIns, df.parity4$DFCFirstHeat, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$DaysToFirstIns, df.parity1$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor(df.parity2$DaysToFirstIns, df.parity2$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor(df.parity3$DaysToFirstIns, df.parity3$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor(df.parity4$DaysToFirstIns, df.parity4$DFCFirstHeat, use = "complete.obs", method = "spearman")

cor.test(df.parity1$DaysToFirstIns, df.parity1$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor.test(df.parity2$DaysToFirstIns, df.parity2$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor.test(df.parity3$DaysToFirstIns, df.parity3$DFCFirstHeat, use = "complete.obs", method = "spearman")
cor.test(df.parity4$DaysToFirstIns, df.parity4$DFCFirstHeat, use = "complete.obs", method = "spearman")
```

```{r}
# Time for DFC_CLA to DFC_FirstHeat

#　
```

# Ketosis parameters
```{r}
## AvgKeto20_atleast10 vs KetoDur20
plot(Three$AvgKeto20_atleast10, Three$KetoDur20, 
     main = "Scatterplot of Avg milk BHB and ketosis duration within 20 DFC", 
     xlab = "Average milk BHB (mmol/L)",
     ylab = "Duraiton (days)", pch = 19)
abline(v = 0.08, col = "red")

## Pearson correlation
cor(Three$AvgKeto20_atleast10, Three$KetoDur20, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(Three$AvgKeto20_atleast10, Three$KetoDur20, use = "complete.obs", method = "spearman")
cor.test(Three$AvgKeto20_atleast10, Three$KetoDur20, use = "complete.obs", method = "spearman")
```
```{r}
## AvgKeto20_atleast10 vs MaxKetoDFC20
plot(Three$MaxKetoDFC20, Three$AvgKeto20_atleast10, 
     main = "Scatterplot of Avg milk BHB and max milk BHB within 20 DFC", 
     xlab = "Maximum milk BHB (mmol/L)",
     ylab = "Average milk BHB (mmol/L)", pch = 19)
abline(v = 0.08, col = "red")

## Pearson correlation
cor(Three$AvgKeto20_atleast10, Three$MaxKetoDFC20, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(Three$AvgKeto20_atleast10, Three$MaxKetoDFC20, use = "complete.obs", method = "spearman")
cor.test(Three$AvgKeto20_atleast10, Three$MaxKetoDFC20, use = "complete.obs", method = "spearman")
```
```{r}
## AvgKeto20_atleast10 vs nHighKeto20
plot( Three$AvgKeto20_atleast10, Three$nHighKeto20,
     main = "Scatterplot of Avg milk BHB and #milk BHB over threshold within 20 DFC", 
     xlab = "Average milk BHB (mmol/L)",
     ylab = "#milk BHB over 0.08 mmol/L", pch = 19)
abline(v = 0.08, col = "red")

## Pearson correlation
cor(Three$AvgKeto20_atleast10, Three$nHighKeto20, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(Three$AvgKeto20_atleast10, Three$nHighKeto20, use = "complete.obs", method = "spearman")
cor.test(Three$AvgKeto20_atleast10, Three$nHighKeto20, use = "complete.obs", method = "spearman")
```
```{r}
## KetoDur20 vs nHighKeto20
plot(Three$KetoDur20, Three$nHighKeto20,
     main = "Scatterplot of ketosis duration and #milk BHB over threshold within 20 DFC", 
     xlab = "Duration (days)",
     ylab = "#milk BHB over 0.08 mmol/L", pch = 19)
abline(v = 0.08, col = "red")

## Pearson correlation
cor(Three$KetoDur20, Three$nHighKeto20, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(Three$KetoDur20, Three$nHighKeto20, use = "complete.obs", method = "spearman")
cor.test(Three$KetoDur20, Three$nHighKeto20, use = "complete.obs", method = "spearman")
```

# Bivariables
## Reproduction + Ketosis parameters - LastAI vs AvgKeto20_atleast10
```{r}
plot(Three$AvgKeto20_atleast10, Three$Last_AI, 
     main = "Scatterplot of number of AI and average milk BHB within 20 DFC",
     xlab = "Average milk BHB within 20 DFC",
     ylab = "#AI", pch = 19)

par(mfrow = c(2, 2))
plot(df.parity1$AvgKeto20_atleast10, df.parity1$Last_AI, 
     #main = "Scatterplot of number of AI and average milk BHB within 20 DFC", 
     ylab = "#AI",
     xlab = "AvgKeto20 (lactation = 1)", pch = 19)
abline(lm(df.parity1$Last_AI~df.parity1$AvgKeto20_atleast10))
abline(v = 0.08, col = "red")

plot(df.parity2$AvgKeto20_atleast10, df.parity2$Last_AI, 
     #main = "Scatterplot of number of AI and average milk BHB within 20 DFC", 
     ylab = "#AI",
     xlab = "AvgKeto20 (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$Last_AI~df.parity2$AvgKeto20_atleast10), col = "red")
abline(v = 0.08)

plot(df.parity3$AvgKeto20_atleast10, df.parity3$Last_AI, 
     #main = "Scatterplot of number of AI and average milk BHB within 20 DFC", 
     ylab = "#AI",
     xlab = "AvgKeto20 (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$Last_AI~df.parity3$AvgKeto20_atleast10), col = "green")
abline(v = 0.08, col = "red")

plot(df.parity4$AvgKeto20_atleast10, df.parity4$Last_AI, 
     #main = "Scatterplot of number of AI and average milk BHB within 20 DFC", 
     ylab = "#AI",
     xlab = "AvgKeto20 (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$Last_AI~df.parity4$AvgKeto20_atleast10), col = "blue")
abline(v = 0.08, col = "red")

## Pearson correlation
cor(df.parity1$AvgKeto20_atleast10, df.parity1$Last_AI, use = "complete.obs", method = "pearson")
cor(df.parity2$AvgKeto20_atleast10, df.parity2$Last_AI, use = "complete.obs", method = "pearson")
cor(df.parity3$AvgKeto20_atleast10, df.parity3$Last_AI, use = "complete.obs", method = "pearson")
cor(df.parity4$AvgKeto20_atleast10, df.parity4$Last_AI, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$AvgKeto20_atleast10, df.parity1$Last_AI, use = "complete.obs", method = "spearman")
cor(df.parity2$AvgKeto20_atleast10, df.parity2$Last_AI, use = "complete.obs", method = "spearman")
cor(df.parity3$AvgKeto20_atleast10, df.parity3$Last_AI, use = "complete.obs", method = "spearman")
cor(df.parity4$AvgKeto20_atleast10, df.parity4$Last_AI, use = "complete.obs", method = "spearman")

cor.test(df.parity1$AvgKeto20_atleast10, df.parity1$Last_AI, use = "complete.obs", method = "spearman")
cor.test(df.parity2$AvgKeto20_atleast10, df.parity2$Last_AI, use = "complete.obs", method = "spearman")
cor.test(df.parity3$AvgKeto20_atleast10, df.parity3$Last_AI, use = "complete.obs", method = "spearman")
cor.test(df.parity4$AvgKeto20_atleast10, df.parity4$Last_AI, use = "complete.obs", method = "spearman")
```

## Reproduction + Ketosis parameters - DFC_CLA vs nHighKeto20
```{r}
plot(Three$nHighKeto20, Three$DFC_CLA, 
     main = "Scatterplot of count milk BHB levels over threshold within 20 DFC and DFC to first cyclicity",
     xlab = "Number of time milk BHB over threshold within 20 DFC",
     ylab = "DFC to first cyclicity", pch = 19)

par(mfrow = c(2, 2))
plot(df.parity1$nHighKeto20, df.parity1$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 1)", pch = 19)
abline(lm(df.parity1$DFC_CLA~df.parity1$nHighKeto20))

plot(df.parity2$nHighKeto20, df.parity2$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DFC_CLA~df.parity2$nHighKeto20), col = "red")

plot(df.parity3$nHighKeto20, df.parity3$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DFC_CLA~df.parity3$nHighKeto20), col = "green")

plot(df.parity4$nHighKeto20, df.parity4$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DFC_CLA~df.parity4$nHighKeto20), col = "blue")

## Pearson correlation
cor(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

cor.test(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")
```

## Reproduction + Ketosis parameters - DFC_CAL vs AvgKeto20_atleast10
```{r}
## DFC_CLA vs Average milk BHB levels
plot(Three$AvgKeto20_atleast10, Three$DFC_CLA, 
     main = "Scatterplot of average milk BHB levels and DFC to first cyclicity",
     xlab = "Average milk BHB levels (mmol/L)",
     ylab = "DFC to first cyclicity", pch = 19)

par(mfrow = c(2, 2))
plot(df.parity1$AvgKeto20_atleast10, df.parity1$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "Average milk BHB levels (mmol/L) (lactation = 1)", pch = 19)
abline(lm(df.parity1$DFC_CLA~df.parity1$AvgKeto20_atleast10))
abline(v = 0.08, col = "red")

plot(df.parity2$AvgKeto20_atleast10, df.parity2$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "Average milk BHB levels (mmol/L) (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DFC_CLA~df.parity2$AvgKeto20_atleast10), col = "red")
abline(v = 0.08)

plot(df.parity3$AvgKeto20_atleast10, df.parity3$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "Average milk BHB levels (mmol/L) (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DFC_CLA~df.parity3$AvgKeto20_atleast10), col = "green")
abline(v = 0.08, col = "red")

plot(df.parity4$AvgKeto20_atleast10, df.parity4$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "Average milk BHB levels (mmol/L) (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DFC_CLA~df.parity4$AvgKeto20_atleast10), col = "blue")
abline(v = 0.08, col = "red")

## Pearson correlation
cor(df.parity1$AvgKeto20_atleast10, df.parity1$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity2$AvgKeto20_atleast10, df.parity2$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity3$AvgKeto20_atleast10, df.parity3$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity4$AvgKeto20_atleast10, df.parity4$DFC_CLA, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$AvgKeto20_atleast10, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity2$AvgKeto20_atleast10, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity3$AvgKeto20_atleast10, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity4$AvgKeto20_atleast10, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

cor.test(df.parity1$AvgKeto20_atleast10, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity2$AvgKeto20_atleast10, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity3$AvgKeto20_atleast10, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity4$AvgKeto20_atleast10, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

```

## Reproduction + Ketosis parameters - DFC_CLA vs nHighKeto20
```{r}
plot(Three$nHighKeto20, Three$DFC_CLA, 
     main = "Scatterplot of count milk BHB levels over threshold within 20 DFC and DFC to first cyclicity",
     xlab = "Number of time milk BHB over threshold within 20 DFC",
     ylab = "DFC to first cyclicity", pch = 19)

par(mfrow = c(2, 2))
plot(df.parity1$nHighKeto20, df.parity1$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 1)", pch = 19)
abline(lm(df.parity1$DFC_CLA~df.parity1$nHighKeto20))

plot(df.parity2$nHighKeto20, df.parity2$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 2)", pch = 19, col = "red")
abline(lm(df.parity2$DFC_CLA~df.parity2$nHighKeto20), col = "red")

plot(df.parity3$nHighKeto20, df.parity3$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 3)", pch = 19, col = "green")
abline(lm(df.parity3$DFC_CLA~df.parity3$nHighKeto20), col = "green")

plot(df.parity4$nHighKeto20, df.parity4$DFC_CLA, 
     #main = "DFC to first cyclicity and average milk BHB within 20 DFC", 
     ylab = "DFC to first cyclicity",
     xlab = "nHighKeto20 (lactation = 4+)", pch = 19, col = "blue")
abline(lm(df.parity4$DFC_CLA~df.parity4$nHighKeto20), col = "blue")

## Pearson correlation
cor(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "pearson")
cor(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")

cor.test(df.parity1$nHighKeto20, df.parity1$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity2$nHighKeto20, df.parity2$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity3$nHighKeto20, df.parity3$DFC_CLA, use = "complete.obs", method = "spearman")
cor.test(df.parity4$nHighKeto20, df.parity4$DFC_CLA, use = "complete.obs", method = "spearman")
```

## Ketosis + MY parameters - AvgKeto20_atleast10 vs AvgMY20
```{r}
plot(Three$AvgKeto20_atleast10, Three$AvgMY20, 
     main = "Scatterplot of average milk yield and DFC to first cyclicity",
     xlab = "Average milk BHB levels within 20 DFC",
     ylab = "Average milk yield in first 20 DFC (L)", pch = 19)

par(mfrow = c(2, 2))
plot(df.parity1$AvgKeto20_atleast10, df.parity1$AvgMY20, 
     #main = "Scatterplot of average milk yield and DFC to first cyclicity", 
     xlab = "AvgKeto20 (lactation = 1)",
     ylab = "AvgMY20", pch = 19)
abline(lm(df.parity1$AvgMY20~df.parity1$AvgKeto20_atleast10))
abline(v = 0.08, col = "red")

plot(df.parity2$AvgKeto20_atleast10, df.parity2$AvgMY20, 
     #main = "Scatterplot of average milk yield and DFC to first cyclicity", 
     xlab = "AvgKeto20 (lactation = 2)",
     ylab = "AvgMY20", pch = 19, col = "red")
abline(lm(df.parity2$AvgMY20~df.parity2$AvgKeto20_atleast10), col = "red")
abline(v = 0.08)

plot(df.parity3$AvgKeto20_atleast10, df.parity3$AvgMY20, 
     #main = "Scatterplot of average milk yield and DFC to first cyclicity", 
     xlab = "AvgKeto20 (lactation = 3)",
     ylab = "AvgMY20", pch = 19, col = "green")
abline(lm(df.parity3$AvgMY20~df.parity3$AvgKeto20_atleast10), col = "green")
abline(v = 0.08, col = "red")

plot(df.parity4$AvgKeto20_atleast10, df.parity4$AvgMY20, 
     #main = "Scatterplot of average milk yield and DFC to first cyclicity", 
     xlab = "AvgKeto20 (lactation = 4+)",
     ylab = "AvgMY20", pch = 19, col = "blue")
abline(lm(df.parity4$AvgMY20~df.parity4$AvgKeto20_atleast10), col = "blue")
abline(v = 0.08, col = "red")

## Pearson correlation
cor(df.parity1$AvgMY20, df.parity1$AvgKeto20_atleast10, use = "complete.obs", method = "pearson")
cor(df.parity2$AvgMY20, df.parity2$AvgKeto20_atleast10, use = "complete.obs", method = "pearson")
cor(df.parity3$AvgMY20, df.parity3$AvgKeto20_atleast10, use = "complete.obs", method = "pearson")
cor(df.parity4$AvgMY20, df.parity4$AvgKeto20_atleast10, use = "complete.obs", method = "pearson")

## Spearman correlation
cor(df.parity1$AvgMY20, df.parity1$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor(df.parity2$AvgMY20, df.parity2$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor(df.parity3$AvgMY20, df.parity3$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor(df.parity4$AvgMY20, df.parity4$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")

## Spearman correlation wtih p-value
cor.test(df.parity1$AvgMY20, df.parity1$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor.test(df.parity2$AvgMY20, df.parity2$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor.test(df.parity3$AvgMY20, df.parity3$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
cor.test(df.parity4$AvgMY20, df.parity4$AvgKeto20_atleast10, use = "complete.obs", method = "spearman")
```
