---
title: "Poisson regression"
author: "Kelly Chin"
date: "4/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R libraries
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(performance)
library(survival)
library(survminer)
library(gtsummary)
library(coxme)
#library(lme4)
#detach("package:lme4", unload=TRUE)
library(nlme)
library(DHARMa)
library(MASS)
library(arm)
library(foreign)
#install.packages("glmmTMB")
library(glmmTMB)
```

```{r}
Three <- read_excel("../data/raw/Three.xlsx")
#Three$CalvingYear <- as.factor(ifelse(Three$CalvingYear == 2018, 1, 2))
summary(Three)

summary(Three$Last_AI)
sd(Three$Last_AI, na.rm = T)
var(Three$Last_AI, na.rm = T)

# Categorize numbers of AI
#Three <- Three %>%
#  mutate(NumAI = ifelse(Three$Last_AI > 4, 4, 
#                        ifelse(Three$Last_AI <= 4 & Three$Last_AI > 2, 3,
#                               ifelse(Three$Last_AI <= 2 & Three$Last_AI > 1, 2, 
#                                      ifelse(Three$Last_AI == 1, 1, NA)))))

Three.offset <- Three %>%
  mutate(Calving.lastAI = difftime(Last_InsDate, CalvingDate, units = "days"))
Three.offset$Calving.lastAI <- as.numeric(round(Three.offset$Calving.lastAI, 0))
summary(Three.offset$Calving.lastAI)

plot(Three.offset$Calving.lastAI, Three.offset$Last_AI)

boxplot(Three.offset$DaysToFirstIns~Three.offset$FarmNo)

summary(Three.offset$Last_AI)
var(Three.offset$Last_AI)
```

# Research Question: what is the assocation of numbers of AI and ketosis?

# AvgKeto20_100
```{r}
p.mod1 <- glm(Last_AI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three)
summary(p.mod1)

qp.mod1 <- glm(Last_AI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = quasipoisson(link = "log"), data = Three)
summary(qp.mod1)

coef1 = coef(p.mod1)
coef2 = coef(qp.mod1)
se.coef1 = se.coef(p.mod1)
se.coef2 = se.coef(qp.mod1)
models.both <- cbind(coef1, se.coef1, coef2, se.coef2, exponent = exp(coef1))
models.both

nb.mod1 <- glm.nb(Last_AI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), data = Three)
summary(nb.mod1)
coef3 = coef(nb.mod1)

####################
##  doesn't work  ##
####################
#nb.mod2 <- glm.nb(Last_AI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason, 
#                  random = ~ 1|FarmNo,
#                  data = Three)

```

## lme with random effect
```{r}
mod1.lme <- lme(Last_AI ~ AvgKeto20_100 + MaxKetoDFC20 + factor(ad_Lactation) + PeakDFC + AvgMY20 + CalvingSeason, 
                random = ~ 1|FarmNo, 
                data = Three, 
                na.action = na.exclude,
                method = "REML")
summary(mod1.lme)

mod2.lme <- lme(Last_AI ~ AvgKeto20_100  + factor(ad_Lactation) + PeakDFC + AvgMY60 + CalvingSeason, 
                random = ~ 1|FarmNo, 
                data = Three, 
                na.action = na.exclude,
                method = "REML")
summary(mod2.lme)
```

## Poisson regression - glm (without random effect)
```{r}
rm(list=ls(pattern="cox."))
boxplot(Three$Last_AI ~ Three$FarmNo)
plot(Three$AvgKeto20_atleast10, Three$Last_AI)
boxplot(Three$AvgKeto20_atleast10 ~ Three$FarmNo)

poi.mod <- glm(Last_AI ~ 1, data = Three, family = poisson(link = "log"))
summary(poi.mod)

ran.poi.mod <- glmer(Last_AI ~ 1 +(1|FarmNo), Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod)

poi.mod1 <- glm(Last_AI ~ AvgKeto20_100, data = Three.offset, family = poisson(link = "log"))
summary(poi.mod1)

ran.poi.mod1 <- glmer(Last_AI ~ AvgKeto20_100 + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod1)

poi.mod2 <- glm(Last_AI ~ AvgKeto20_100 + factor(ad_Lactation), data = Three.offset, family = poisson(link = "log"))
summary(poi.mod2)
```

## Quasi-poisson regression - glm (without random effect)



## Offset - farmsize
```{r}
mod1 <- glm(Last_AI ~ AvgKeto20_100, data = Three, family = poisson(link = "log"))
summary(mod1)

#mod2 <- glm(Last_AI ~ AvgKeto20_100 + offset(log(CowNo)), data = Three, family = poisson(link = "log"))
#(mod2)
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

mod3 <- glmer(Last_AI ~ AvgKeto20_100 + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(mod3)
res3 <- simulateResiduals(mod3)
plot(res3)

#mod4 <- glmer(Last_AI ~ AvgKeto20_100 + offset(log(CowNo)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

mod4 <- glmer(Last_AI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
# This model estimates the mean numbers of AI within the estimated observation time (per month), with fix average ketosis and fix farm effect (estimate the effect of farm size)
summary(mod4)
res4 <- simulateResiduals(mod4)
plot(res4)

mod5 <- glmer(Last_AI ~ AvgKeto20_100 + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod5)


summary(mod4)
res4 <- simulateResiduals(mod4)
plot(res4)

boxplot(Three$AvgKeto20_100 ~ Three$Last_AI)
boxplot(Three$Last_AI ~ Three$FarmNo)
boxplot(Three$AvgKeto20_100 ~ Three$FarmNo)

ggplot(data = Three, aes(x = Last_AI, y = AvgKeto20_100)) +
  geom_bar(stat = "identity")
```

## glmer with offset (time from calving to last AI)
```{r}
mean(Three.offset$Calving.lastAI, na.rm = TRUE)
boxplot(Three.offset$Calving.lastAI~Three.offset$Last_AI)
Three.offset$Last_AI <- as.factor(Three.offset$Last_AI)
box_plot <- ggplot(data = Three.offset, aes(x = Last_AI, y = Calving.lastAI)) +
  geom_boxplot() +
  geom_jitter(shape = 15, 
              color = "steelblue", 
              position = position_jitter(width = 0.15)) +
  xlab("numbers of AI") + ylab("time from calving to last AI") +
  theme_classic()
box_plot


mod1.offset <- glmer(Last_AI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.offset)
res.mod1.offset <- simulateResiduals(mod1.offset)
plot(res.mod1.offset)
```https://stats.stackexchange.com/questions/11182/when-to-use-an-offset-in-a-poisson-regression

# glmer without offset
```{r}
Three.offset$Last_AI <- as.numeric(Three.offset$Last_AI)
mod1.glmer <- glmer(Last_AI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer)

?I
plot(Three.offset$PeakDFC ~ Three.offset$Last_AI)
summary(Three.offset$AvgKeto20_100)
summary(Three.offset$PeakDFC)
summary(Three.offset$AvgMY60)
hist(Three.offset$AvgMY60)
cor(Three.offset$AvgMY60, Three.offset$PeakDFC, use = "complete.obs") # -0.29
cor(Three.offset$ad_Lactation, Three.offset$AvgMY60, use = "complete.obs") # 0.52
cor(Three.offset$AvgMY60, Three.offset$AvgKeto20_100, use = "complete.obs") # 0.15

prop.table(table(Three.offset$ad_Lactation, cut(Three.offset$AvgMY60, c(0, 30, 37, 44, 70))), 1)
prop.table(table(Three.offset$ad_Lactation, Three.offset$CalvingSeason))

mod2.glmer <- glmer(Last_AI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + cut(PeakDFC, c(0, 30, 60, 90, 305)) + I(AvgMY60/5) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
# doesn't work

mod1 <- glmer(Last_AI ~ AvgKeto20_100 + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1)
mod2 <- glmer(Last_AI ~ factor(ad_Lactation) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2)
mod3 <- glmer(Last_AI ~ CalvingSeason + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3)
mod4 <- glmer(Last_AI ~ I(PeakDFC/30) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod4)
mod5 <- glmer(Last_AI ~  cut(AvgMY60, c(0, 30, 37, 44, 70)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod5)
mod6 <- glmer(Last_AI ~  cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

mod2.glmer <- glmer(Last_AI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer)
res.mod2.glmer <- simulateResiduals(mod2.glmer)
plot(res.mod2.glmer)

########################################
##  lactation as continuous variable  ##
########################################
mod1.glmer1 <- glmer(Last_AI ~ AvgKeto20_100 + Lactation + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer1)

mod2.glmer1 <- glmer(Last_AI ~ AvgKeto20_100  + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer1)

mod3.glmer1 <- glmer(Last_AI ~ CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3.glmer1)
#############################################
#############################################

# -AvgKeto20_100
mod2.glmer <- glmer(Last_AI ~ factor(ad_Lactation) + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer)

# -CowNo (-farm size)
mod3.glmer <-  glmer(Last_AI ~ factor(ad_Lactation) + CalvingSeason + PeakDFC + AvgMY60  + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3.glmer)

summary(Three.offset$Last_AI)
```

## glmmTMB
```{r}
fit1 <- glmmTMB(Last_AI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(fit1)
#res.fit1 <- simulateResiduals(fit1)
#plot(res.fit1)

fit1.nb <- glmmTMB(Last_AI ~ AvgKeto20_100 + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = nbinom1)

poisson1 <- glmer(Last_AI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(poisson1)
```

## Offset - time from calving to last AI
```{r}
mod5 <- glm(Last_AI ~ AvgKeto20_100 + offset(log(Calving.lastAI)), data = Three.offset, family = poisson(link = "log"))
summary(mod5)

mod6 <- glmer(Last_AI ~ AvgKeto20_100 + offset(log(Calving.lastAI)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

summary(mod6)
res6 <- simulateResiduals(mod6)
plot(res6)

anova(mod5, mod6)
drop1(mod6)
```

## glmer
```{r}
mod1.poisson <- glmer(Last_AI ~ AvgKeto20_atleast10 + MaxKetoDFC20 + factor(ad_Lactation) + PeakDFC + AvgMY20 + CalvingSeason + (1 | FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(mod1.poisson)

mod0.poisson <- glmer(Last_AI ~ 1 + (1 | FarmNo), data = Three, family = poisson(link = "log"))
summary(mod2.poisson)
```

## glmer 
```{r}
poi.avgketo20 <- glmer(Last_AI~ AvgKeto20_atleast10 + (1 | FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20)
anova(poi.mod, poi.avgketo20, test = "Chisq")
drop1(poi.avgketo20)

poi.avgketo20.1 <- glmer(Last_AI~ AvgKeto20_atleast10 + (FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.1)

poi.avgketo20.cate <- glmer(Last_AI~ cut(AvgKeto20_atleast10, c(0, 0.04, 0.07, 0.1, 0.5)) + (1 | FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.cate)

poi.avgketo20.100 <- glmer(Last_AI~ AvgKeto20_100 + (1 | FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.100)

```

# KetoDur20
```{r}
rm(list=ls(pattern="mod"))
summary(Three$KetoDur20)
#Days of ketosis
Three$KetoDur20 <- as.factor(Three$KetoDur20)
ggplot(data=Three, aes(x=factor(KetoDur20)))+
  geom_bar(stat="count", fill="lightblue") +
  stat_count(geom="text", size=3, aes(label=..count..)) +
  ggtitle("Distribution of KetoDur20") +
  theme_classic() +
  xlab("days of milk BHB over 0.08 mmol/L within 20 DFC") +
  ylab("#cows")

Three$KetoDur20 <- as.numeric(Three$KetoDur20)
summary(Three$Last_AI)
boxplot(Three$Last_AI ~ Three$FarmNo)
boxplot(Three$KetoDur20 ~ Three$Last_AI)
boxplot(Three$KetoDur20 ~ Three$FarmNo)
```

## glm
```{r}
poi.mod <- glm(Last_AI ~ 1, data = Three, family = poisson(link = "log"))
summary(poi.mod)

poi.mod1 <- glm(Last_AI ~ KetoDur20, data = Three, family = poisson(link = "log"))
summary(poi.mod1)

p.mod <- glm(Last_AI ~ KetoDur20 + factor(ad_Lactation) + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three)
summary(p.mod)
#plot(p.mod)

p.mod_no_Lact <- glm(Last_AI ~ KetoDur20 + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three) 
summary(p.mod_no_Lact)
#plot(p.mod_no_Lact)
```

## Offset - farmsize
```{r}
mod1 <- glm(Last_AI ~ KetoDur20, data = Three, family = poisson(link = "log"))
summary(mod1)

#mod2 <- glm(Last_AI ~ KetoDur20 + offset(log(CowNo)), data = Three, family = poisson(link = "log"))
#summary(mod2)
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size

mod3 <- glmer(Last_AI ~ KetoDur20 + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(mod3)

#mod4 <- glmer(Last_AI ~ KetoDur20 + offset(log(CowNo)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

anova(mod1, mod2, test = "Chisq")
anova(mod1, mod3, test = "Chisq")
plot(mod3)
anova(mod3, mod4, test="Chisq")

res3 <- simulateResiduals(mod3)
plot(res3)
```

## Offset - time from calving to last AI
```{r}
mod5 <- glm(Last_AI ~ KetoDur20 + offset(log(Calving.lastAI)), data = Three.offset, family = poisson(link = "log"))
summary(mod5)

mod6 <- glmer(Last_AI ~ KetoDur20 + offset(log(Calving.lastAI)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

anova(mod5, mod6)
drop1(mod6)
```

# Extra AI
```{r}
extra <- Three.offset %>%
  filter(Last_AI >= 1)
mean(extra$Last_AI)
var(extra$Last_AI)
table(extra$Last_AI)

mod.extra <- glmer(Last_AI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = extra, family = poisson(link = "log"))
summary(mod.extra)
res.extra <- simulateResiduals(mod.extra)
plot(res.extra)

mod.extra2 <- glmer(Last_AI ~ AvgKeto20_100)


```
