---
title: "Poisson regression"
author: "Kelly Chin"
date: "4/17/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R libraries
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(performance)
library(survival)
library(survminer)
library(gtsummary)
library(coxme)
#library(lme4)
#detach("package:lme4", unload=TRUE)
library(nlme)
library(DHARMa)
library(MASS)
library(arm)
library(foreign)
#install.packages("glmmTMB")
library(glmmTMB)
library(car)
```

```{r}
Three <- read_excel("../data/raw/Three.xlsx")
#Three$CalvingYear <- as.factor(ifelse(Three$CalvingYear == 2018, 1, 2))
summary(Three)

summary(Three$NumAI)
sd(Three$NumAI, na.rm = T)
var(Three$NumAI, na.rm = T)

# Categorize numbers of AI
#Three <- Three %>%
#  mutate(NumAI = ifelse(Three$NumAI > 4, 4, 
#                        ifelse(Three$NumAI <= 4 & Three$NumAI > 2, 3,
#                               ifelse(Three$NumAI <= 2 & Three$NumAI > 1, 2, 
#                                      ifelse(Three$NumAI == 1, 1, NA)))))

Three.offset <- Three %>%
  mutate(Calving.lastAI = difftime(Last_InsDate, CalvingDate, units = "days"))
Three.offset$Calving.lastAI <- as.numeric(round(Three.offset$Calving.lastAI, 0))
summary(Three.offset$Calving.lastAI)

plot(Three.offset$Calving.lastAI, Three.offset$NumAI)

boxplot(Three.offset$DaysToFirstIns~Three.offset$FarmNo)

summary(Three.offset$NumAI)
var(Three.offset$NumAI)

hist(Three$MaxKeto)
```

# Research Question: what is the assocation of numbers of AI and ketosis?

# AvgKeto20_100
```{r}
p.mod1 <- glm(NumAI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three)
summary(p.mod1)

qp.mod1 <- glm(NumAI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = quasipoisson(link = "log"), data = Three)
summary(qp.mod1)

coef1 = coef(p.mod1)
coef2 = coef(qp.mod1)
se.coef1 = se.coef(p.mod1)
se.coef2 = se.coef(qp.mod1)
models.both <- cbind(coef1, se.coef1, coef2, se.coef2, exponent = exp(coef1))
models.both

nb.mod1 <- glm.nb(NumAI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), data = Three)
summary(nb.mod1)
coef3 = coef(nb.mod1)

####################
##  doesn't work  ##
####################
#nb.mod2 <- glm.nb(NumAI ~ AvgKeto20_atleast10 + ad_Lactation + PeakDFC + AvgMY60 + CalvingSeason, 
#                  random = ~ 1|FarmNo,
#                  data = Three)

```

## lme with random effect
```{r}
mod1.lme <- lme(NumAI ~ AvgKeto20_100 + MaxKetoDFC20 + factor(ad_Lactation) + PeakDFC + AvgMY20 + CalvingSeason, 
                random = ~ 1|FarmNo, 
                data = Three, 
                na.action = na.exclude,
                method = "REML")
summary(mod1.lme)

mod2.lme <- lme(NumAI ~ AvgKeto20_100  + factor(ad_Lactation) + PeakDFC + AvgMY60 + CalvingSeason, 
                random = ~ 1|FarmNo, 
                data = Three, 
                na.action = na.exclude,
                method = "REML")
summary(mod2.lme)
```

## Poisson regression & Quasi-poisson regression - glm (without random effect)
```{r}
rm(list=ls(pattern="cox."))
boxplot(Three$NumAI ~ Three$FarmNo)
plot(Three$AvgKeto20_atleast10, Three$NumAI)
boxplot(Three$AvgKeto20_atleast10 ~ Three$FarmNo)

qpoi.mod <- glm(NumAI ~ 1, data = Three, family = quasipoisson(link = "log"))
summary(qpoi.mod)
poi.mod <- glm(NumAI ~ 1, data = Three, family = poisson(link = "log"))
summary(poi.mod)
ran.poi.mod <- glmer(NumAI ~ 1 +(1|FarmNo), Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod)

poi.mod1 <- glm(NumAI ~ AvgKeto20_100, data = Three.offset, family = poisson(link = "log"))
summary(poi.mod1)
qpoi.mod1 <- glm(NumAI ~ AvgKeto20_100, data = Three.offset, family = quasipoisson(link = "log"))
summary(qpoi.mod1)
ran.poi.mod1 <- glmer(NumAI ~ AvgKeto20_100 + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod1)

poi.mod2 <- glm(NumAI ~ AvgKeto20_100 + factor(ad_Lactation), data = Three.offset, family = poisson(link = "log"))
summary(poi.mod2)
## to find out the outliers 
plot(poi.mod2) 
## to find out the influential observations
influencePlot(poi.mod2)

qpoi.mod2 <- glm(NumAI ~ AvgKeto20_100 + factor(ad_Lactation), data = Three.offset, family = quasipoisson(link = "log"))
summary(qpoi.mod2)
# outliers
plot(qpoi.mod2)
# influential observations
influencePlot(qpoi.mod2)


ran.poi.mod2 <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) +(1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod2)

ran.poi.mod3 <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(ran.poi.mod3)
```

## Offset - farmsize
```{r}
mod1 <- glm(NumAI ~ AvgKeto20_100, data = Three, family = poisson(link = "log"))
summary(mod1)

#mod2 <- glm(NumAI ~ AvgKeto20_100 + offset(log(CowNo)), data = Three, family = poisson(link = "log"))
#(mod2)
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

mod3 <- glmer(NumAI ~ AvgKeto20_100 + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(mod3)
res3 <- simulateResiduals(mod3)
plot(res3)

#mod4 <- glmer(NumAI ~ AvgKeto20_100 + offset(log(CowNo)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

mod4 <- glmer(NumAI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
# This model estimates the mean numbers of AI within the estimated observation time (per month), with fix average ketosis and fix farm effect (estimate the effect of farm size)
summary(mod4)
res4 <- simulateResiduals(mod4)
plot(res4)

mod5 <- glmer(NumAI ~ AvgKeto20_100 + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod5)


summary(mod4)
res4 <- simulateResiduals(mod4)
plot(res4)

boxplot(Three$AvgKeto20_100 ~ Three$NumAI)
boxplot(Three$NumAI ~ Three$FarmNo)
boxplot(Three$AvgKeto20_100 ~ Three$FarmNo)

ggplot(data = Three, aes(x = NumAI, y = AvgKeto20_100)) +
  geom_bar(stat = "identity")
```

## glmer with offset (time from calving to last AI)
```{r}
mean(Three.offset$Calving.lastAI, na.rm = TRUE)
boxplot(Three.offset$Calving.lastAI~Three.offset$NumAI)
Three.offset$NumAI <- as.factor(Three.offset$NumAI)
box_plot <- ggplot(data = Three.offset, aes(x = NumAI, y = Calving.lastAI)) +
  geom_boxplot() +
  geom_jitter(shape = 15, 
              color = "steelblue", 
              position = position_jitter(width = 0.15)) +
  xlab("numbers of AI") + ylab("time from calving to last AI") +
  theme_classic()
box_plot


mod1.offset <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.offset)
res.mod1.offset <- simulateResiduals(mod1.offset)
plot(res.mod1.offset)

offset1 <- glmer(NumAI ~ AvgKeto20_100 + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset1)
offset1.1 <- glmer(NumAI ~ AvgKeto20_100 + cut(AvgMY60, c(0, 30, 37, 44, 70)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset1.1)
offset1.2 <- glmer(NumAI ~ AvgKeto20_100 + I(PeakDFC/30) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset1.2)


offset2 <- glmer(NumAI ~ factor(ad_Lactation) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset2)
offset3 <- glmer(NumAI ~ CalvingSeason + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset3)
offset4 <- glmer(NumAI ~ I(PeakDFC/30) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset4)
offset5 <- glmer(NumAI ~ cut(AvgMY60, c(0, 30, 37, 44, 70)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset5)
offset6 <- glmer(NumAI ~ cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(offset6)

```

# glmer without offset
```{r}
Three.offset$NumAI <- as.numeric(Three.offset$NumAI)

# Milk - PeakDFC, AvgMY60
mod1.glmer <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer) # AIC = 15996.6

mod1.glmer1 <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))

# Milk - PeakMY
mod1.glmer.my <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + cut(PeakMY, c(0, 43, 52, 60, 110)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer.my) # AIC = 16008.8

# Milk - PeakMY, PeakDFC
mod1.glmer.my.dfc <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason +  I(PeakDFC/30) + cut(PeakMY, c(0, 43, 52, 60, 110)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer.my.dfc) # AIC = 16006.7

# without lactation groups
mod1.glmer.nolact <-  glmer(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer.nolact) # AIC = 15994.4
res.mod1.nolact <- simulateResiduals(mod1.glmer.nolact)
plot(res.mod1.nolact)
plot(mod1.glmer.nolact)

?I
plot(Three.offset$PeakDFC ~ Three.offset$NumAI)
summary(Three.offset$AvgKeto20_100)
summary(Three.offset$PeakDFC)
summary(Three.offset$AvgMY60)
summary(Three.offset$PeakMY)
hist(Three.offset$AvgMY60)
cor(Three.offset$AvgMY60, Three.offset$PeakDFC, use = "complete.obs") # -0.29
cor(Three.offset$ad_Lactation, Three.offset$AvgMY60, use = "complete.obs") # 0.52
cor(Three.offset$AvgMY60, Three.offset$AvgKeto20_100, use = "complete.obs") # 0.15

prop.table(table(Three.offset$ad_Lactation, cut(Three.offset$AvgMY60, c(0, 30, 37, 44, 70))), 1)
prop.table(table(Three.offset$ad_Lactation, Three.offset$CalvingSeason))

mod2.glmer <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + cut(PeakDFC, c(0, 30, 60, 90, 305)) + I(AvgMY60/5) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
# doesn't work

mod1 <- glmer(NumAI ~ AvgKeto20_100 + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1)
mod2 <- glmer(NumAI ~ factor(ad_Lactation) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2)
mod3 <- glmer(NumAI ~ CalvingSeason + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3)
mod4 <- glmer(NumAI ~ I(PeakDFC/7) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod4)
mod5 <- glmer(NumAI ~  cut(AvgMY60, c(0, 30, 37, 44, 70)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod5)
mod6 <- glmer(NumAI ~  cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

# without farm size
mod2.glmer <- glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer) # AIC = 15998.1
res.mod2.glmer <- simulateResiduals(mod2.glmer)
plot(res.mod2.glmer) 
exp(0.004767)
exp(-0.070481)

########################################
##  lactation as continuous variable  ##
########################################
mod1.glmer1 <- glmer(NumAI ~ AvgKeto20_100 + Lactation + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer1)

mod2.glmer1 <- glmer(NumAI ~ AvgKeto20_100  + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer1)

mod3.glmer1 <- glmer(NumAI ~ CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3.glmer1)
#############################################
#############################################

# -AvgKeto20_100
mod2.glmer <- glmer(NumAI ~ factor(ad_Lactation) + CalvingSeason + PeakDFC + AvgMY60 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod2.glmer)

# -CowNo (-farm size)
mod3.glmer <-  glmer(NumAI ~ factor(ad_Lactation) + CalvingSeason + PeakDFC + AvgMY60  + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod3.glmer)

summary(Three.offset$NumAI)

########################
##     Interaction    ##
########################
mod1.int <-  glmer(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + factor(ad_Lactation):AvgKeto20_100 + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
# doesn't work
```

## glmmTMB (negative binominal with random effect)
```{r}
fit1 <- glmmTMB(NumAI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit1)
res.fit1 <- simulateResiduals(fit1)
plot(res.fit1)

fit1.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + (1|FarmNo), data = Three, family = nbinom1) #AIC 15827
fit1.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + (1|FarmNo), data = Three, family = nbinom2) #AIC 15791
anova(fit1.nb, fit1.nb1)
summary(fit1.nb1)

fit2.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + (1|FarmNo), data = Three, family = nbinom1)
fit2.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + (1|FarmNo), data = Three, family = nbinom2)
summary(fit2.nb) #AIC 15820.6
summary(fit2.nb1) #AIC 15783.7

fit3.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + (1|FarmNo), data = Three, family = nbinom1)
fit3.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + (1|FarmNo), data = Three, family = nbinom2)
summary(fit3.nb) #AIC 15813.7
summary(fit3.nb1) #AIC 15778.6

#fit4.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + AvgMY60 + (1|FarmNo), data = Three, family = nbinom1)
fit4.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + AvgMY60 + (1|FarmNo), data = Three, family = nbinom2)
#summary(fit4.nb) #AIC 15754.3
summary(fit4.nb1) #AIC 15718.1
res.fit4 <- simulateResiduals(fit4.nb1)
plot(res.fit4)

#fit5.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + PeakMY + AvgMY60 + (1|FarmNo), data = Three, family = nbinom1)
fit5.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + PeakMY + AvgMY60 + (1|FarmNo), data = Three, family = nbinom2)
#summary(fit5.nb) # AIC 15756.1, PeakMY IS NOT SIGNIFICANT
summary(fit5.nb1) # AIC 15719.1, PeakMY IS NOT SIGNIFICANT

####################################################
##  PeakDFC and AvgMY60 are the best combination  ##
####################################################

#fit6.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + (1|FarmNo), data = Three, family = nbinom1)
fit6.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + (1|FarmNo), data = Three, family = nbinom2)
fit6.nb3 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + (1|FarmNo), data = Three, family = poisson(link = "log"))
fit6.nb4 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60, data = Three, family = poisson(link = "log"))
#summary(fit6.nb) #AIC 15749.2
summary(fit6.nb1) #AIC 15712.7
summary(fit6.nb3) #AIC 15976.9
summary(fit6.nb4) #AIC 16433.5, model without random effect is no better than the model with random effect, so random intercept is taking a big part in the model
anova(fit6.nb1, fit6.nb3)

res.fit6 <- simulateResiduals(fit6.nb1)
plot(res.fit6)
residuals(res.fit6, quantileFunction = qnorm, outlierValues = c(-2, 2))

########################
#  Poisson regression  #
########################

mod1.glmer.nolact <-  glmer(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod1.glmer.nolact) #AIC 15994.4

plot(mod1.glmer.nolact)

###############################
##  use PeakMY (not AvgMY)   ##
###############################

#fit7.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + PeakMY + (1|FarmNo), data = Three, family = nbinom1)
fit7.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + PeakMY + (1|FarmNo), data = Three, family = nbinom2)
fit7.poisson <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + PeakMY + (1|FarmNo), data = Three, family = poisson(link = "log"))
#summary(fit7.nb) #AIC 15767.6, Lactation, PeakDFC all become not significant
summary(fit7.nb1) #AIC 15727.3, Lactation, CalvingSeason, PeakDFC all become not significant
summary(fit7.poisson) #AIC 15999.1, Lactation is not significant
res.fit7 <- simulateResiduals(fit7.nb1)
plot(res.fit7)

fit7.poisson2 <- glmmTMB(NumAI ~ AvgKeto20_100 + MaxKetoDFC60 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + PeakMY + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit7.poisson2) #AIC 15986.7, Lactation is not significant

########################################
##  AvgMY60 is a continuous variable  ##
########################################

#fit8.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500))  + (1|FarmNo), data = Three, family = nbinom1)
fit8.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = nbinom2)
fit8.nb3 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
#summary(fit8.nb) #AIC 15747.1
summary(fit8.nb1) #AIC 15711.3
summary(fit8.nb3) #AIC 15975.6
anova(mod1.glmer.nolact, fit8.nb1, test = "Chisq")

res.fit8 <- simulateResiduals(fit8.nb1)
plot(res.fit8)
boxplot(residuals(fit8.nb1, "deviance"))$out  #doesn't work
max(resid(fit8.nb1, "deviance"))$out          #doesn't work

#########################
##  without lactation  ##
#########################
fit8.nb2 <- glmmTMB(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = nbinom2)
summary(fit8.nb2) #AIC 15710.0
res.fit8.nb2 <- simulateResiduals(fit8.nb2)
plot(res.fit8.nb2)

fit8.poisson <- glmmTMB(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit8.poisson) #AIC 15976.2
anova(fit8.nb2, fit8.poisson)
res.fit8.poisson <- simulateResiduals(fit8.poisson)
plot(res.fit8.poisson)

# add MaxKetoDFC60 (no lactation)
fit8.poisson2 <- glmmTMB(NumAI ~ AvgKeto20_100 + MaxKetoDFC60 + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit8.poisson2) #AIC 15963.7
anova(fit8.poisson, fit8.poisson2)

# add Lactation (has everything)
fit8.poisson3 <- glmmTMB(NumAI ~ AvgKeto20_100 + MaxKetoDFC60 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit8.poisson3) #AIC 15962.9

anova(fit8.poisson2, fit8.poisson3) # model fit8.poisson2 and fit8.poisson3 are similar
anova(fit7.poisson2, fit8.poisson3) # model fit8.poisson3 is better than model fit7.poisson (AvgMY60 is better than PeakMY)

# remove CalvingSeason (has lactation)
fit8.poisson4 <- glmmTMB(NumAI ~ AvgKeto20_100 + MaxKetoDFC60 + factor(ad_Lactation) + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit8.poisson4) #AIC 15967.7
anova(fit8.poisson2, fit8.poisson4) # model fit8.poisson2 and fit8.poisson4 are similar
anova(fit8.poisson3, fit8.poisson4) # model fit8.poisson3 is slightly better than model fit8.poisson4

# lactation become continuous variable -> less significant
fit8.poisson5 <- glmmTMB(NumAI ~ AvgKeto20_100 + MaxKetoDFC60 + Lactation + I(PeakDFC/7) + AvgMY60 + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit8.poisson5)

## TODO
### consider not to have lactation in my model!!

###########################
##  Categorized AvgMY60  ##
###########################

#fit9.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500))  + (1|FarmNo), data = Three, family = nbinom1)
fit9.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = nbinom2)
#summary(fit9.nb) #AIC 15762.5, Lactation is no longer significant!!
summary(fit9.nb1) #AIC 15727.2, Lactation is no longer significant!!

fit9.poisson <- glmmTMB(NumAI ~ AvgKeto20_100 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit9.poisson) #AIC 15996.6, NO BETTER THAN MODEL fit8.poisson

#######################################################
##  Categorized AvgMY60 and model without Lactation  ##
#######################################################

#fit10.nb <- glmmTMB(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500))  + (1|FarmNo), data = Three, family = nbinom1)
fit10.nb1 <- glmmTMB(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = nbinom2)
#summary(fit10.nb) #AIC 15759.2
summary(fit10.nb1) #AIC 15723.8
res.fit10 <- simulateResiduals(fit10.nb1)
plot(res.fit10)

fit10.poisson <- glmmTMB(NumAI ~ AvgKeto20_100 + CalvingSeason + I(PeakDFC/7) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 500)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(fit10.poisson) #AIC 15994.4, NO BETTER THAN MODEL fit9.poisson

######################################
##  fit8.poisson2 is the best model  ##
######################################
res.fit8.poisson2 <- simulateResiduals(fit8.poisson2)
plot(res.fit8.poisson2)

```

## Offset - time from calving to last AI
```{r}
mod5 <- glm(NumAI ~ AvgKeto20_100 + offset(log(Calving.lastAI)), data = Three.offset, family = poisson(link = "log"))
summary(mod5)

mod6 <- glmer(NumAI ~ AvgKeto20_100 + offset(log(Calving.lastAI)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

summary(mod6)
res6 <- simulateResiduals(mod6)
plot(res6)

anova(mod5, mod6)
drop1(mod6)
```

## glmer 
```{r}
poi.avgketo20 <- glmer(NumAI~ AvgKeto20_atleast10 + (1|FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20)
anova(poi.mod, poi.avgketo20, test = "Chisq")
drop1(poi.avgketo20)

poi.avgketo20.1 <- glmer(NumAI~ AvgKeto20_atleast10 + (FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.1)

poi.avgketo20.cate <- glmer(NumAI~ cut(AvgKeto20_atleast10, c(0, 0.04, 0.07, 0.1, 0.5)) + (1 | FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.cate)

poi.avgketo20.100 <- glmer(NumAI~ AvgKeto20_100 + (1|FarmNo), 
                      data = Three, family = poisson(link = "log"))
summary(poi.avgketo20.100)

```

# KetoDur20
```{r}
rm(list=ls(pattern="mod"))
summary(Three$KetoDur20)
#Days of ketosis
Three$KetoDur20 <- as.factor(Three$KetoDur20)
ggplot(data=Three, aes(x=factor(KetoDur20)))+
  geom_bar(stat="count", fill="lightblue") +
  stat_count(geom="text", size=3, aes(label=..count..)) +
  ggtitle("Distribution of KetoDur20") +
  theme_classic() +
  xlab("days of milk BHB over 0.08 mmol/L within 20 DFC") +
  ylab("#cows")

Three$KetoDur20 <- as.numeric(Three$KetoDur20)
summary(Three$NumAI)
boxplot(Three$NumAI ~ Three$FarmNo)
boxplot(Three$KetoDur20 ~ Three$NumAI)
boxplot(Three$KetoDur20 ~ Three$FarmNo)
```

## glm
```{r}
poi.mod <- glm(NumAI ~ 1, data = Three, family = poisson(link = "log"))
summary(poi.mod)

poi.mod1 <- glm(NumAI ~ KetoDur20, data = Three, family = poisson(link = "log"))
summary(poi.mod1)

p.mod <- glm(NumAI ~ KetoDur20 + factor(ad_Lactation) + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three)
summary(p.mod)
#plot(p.mod)

p.mod_no_Lact <- glm(NumAI ~ KetoDur20 + PeakDFC + AvgMY60 + CalvingSeason + factor(FarmNo), family = poisson(link = "log"), data = Three) 
summary(p.mod_no_Lact)
#plot(p.mod_no_Lact)
```

## Offset - farmsize
```{r}
mod1 <- glm(NumAI ~ KetoDur20, data = Three, family = poisson(link = "log"))
summary(mod1)

#mod2 <- glm(NumAI ~ KetoDur20 + offset(log(CowNo)), data = Three, family = poisson(link = "log"))
#summary(mod2)
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size

mod3 <- glmer(NumAI ~ KetoDur20 + (1|FarmNo), data = Three, family = poisson(link = "log"))
summary(mod3)

#mod4 <- glmer(NumAI ~ KetoDur20 + offset(log(CowNo)) + (1|FarmNo), data = Three, family = poisson(link = "log"))
## This model doesn't mean anything because the outcome is the 'mean numbers of insemination of an individual cow' and should not be divided to farm size -> shouldn't use farm size in offset! 

anova(mod1, mod2, test = "Chisq")
anova(mod1, mod3, test = "Chisq")
plot(mod3)
anova(mod3, mod4, test="Chisq")

res3 <- simulateResiduals(mod3)
plot(res3)
```

## Offset - time from calving to last AI
```{r}
mod5 <- glm(NumAI ~ KetoDur20 + offset(log(Calving.lastAI)), data = Three.offset, family = poisson(link = "log"))
summary(mod5)

mod6 <- glmer(NumAI ~ KetoDur20 + offset(log(Calving.lastAI)) + (1|FarmNo), data = Three.offset, family = poisson(link = "log"))
summary(mod6)

anova(mod5, mod6)
drop1(mod6)
```

# nHighKeto20
```{r}
#numbers of BHB over 0.08 mmol/L within 20 DFC
Three$nHighKeto20 <- as.factor(Three$nHighKeto20)
ggplot(data=Three, aes(x=factor(nHighKeto20)))+
  geom_bar(stat="count", fill="lightblue") +
  stat_count(geom="text", size=3, aes(label=..count..)) +
  ggtitle("Distribution of nHighKeto20") +
  theme_classic() +
  xlab("numbers of milk BHB over 0.08 mmol/L within 20 DFC") +
  ylab("#cows")

Three$nHighKeto20 <- as.numeric(Three$nHighKeto20)
summary(Three$nHighKeto20)

Three.count <- Three %>%
  mutate(HighKeto20 = ifelse(Three$nHighKeto20 > 0, "1", Three$nHighKeto20))
summary(as.factor(Three.count$HighKeto20))
```

## discriptive + model
```{r}
Three.count$HighKeto20 <- as.factor(Three.count$HighKeto20)
Three.count$NumAI <- as.numeric(Three.count$NumAI)
count1.glmer <- glmer(NumAI ~ HighKeto20 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) + cut(CowNo, c(0, 100, 180, 1000)) + (1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count1.glmer)

count2.glmer <- glmer(NumAI ~ HighKeto20 + factor(ad_Lactation) + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) +(1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count2.glmer)

count3.glmer <- glmer(NumAI ~ HighKeto20 + CalvingSeason + I(PeakDFC/30) + cut(AvgMY60, c(0, 30, 37, 44, 70)) +(1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count3.glmer)
res3 <- simulateResiduals(count3.glmer)
plot(res3)

?I
plot(Three.count$HighKeto20, Three.count$NumAI)
summary(Three.count$nHighKeto20)
summary(Three.count$PeakDFC)
summary(Three.count$AvgMY60)
hist(Three.count$AvgMY60)
cor(Three.count$AvgMY60, Three.count$PeakDFC, use = "complete.obs") # -0.29
cor(Three.count$ad_Lactation, Three.count$AvgMY60, use = "complete.obs") # 0.52
cor(Three.count$AvgMY60, Three.count$nHighKeto20, use = "complete.obs") # 0.15

prop.table(table(Three.count$ad_Lactation, cut(Three.count$AvgMY60, c(0, 30, 37, 44, 70))), 1)
prop.table(table(Three.count$ad_Lactation, Three.count$HighKeto20))
cor(Three$ad_Lactation, Three$nHighKeto20)
cor(Three$NumAI, Three$nHighKeto20)

count1 <- glmer(NumAI ~ HighKeto20 + (1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count1)
count2 <- glmer(NumAI ~ HighKeto20 + factor(ad_Lactation) + (1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count2)
count3 <- glmer(NumAI ~ HighKeto20 + factor(ad_Lactation) + CalvingSeason + (1|FarmNo), data = Three.count, family = poisson(link = "log"))
summary(count3)
```


# Extra AI
```{r}
extra <- Three.offset %>%
  filter(NumAI >= 1)
mean(extra$NumAI)
var(extra$NumAI)
table(extra$NumAI)

mod.extra <- glmer(NumAI ~ AvgKeto20_100 + cut(CowNo, c(0, 100, 180, 1000)) + offset(log(Calving.lastAI/30)) + (1|FarmNo), data = extra, family = poisson(link = "log"))
summary(mod.extra)
res.extra <- simulateResiduals(mod.extra)
plot(res.extra)

mod.extra2 <- glmer(NumAI ~ AvgKeto20_100)
```
